%% Plato FSM Translation Tech Memo
%% FSM Translation Algorithm
%%
%% Adrian Wheeldon
%% March 2017
\section{Concepts to FSM translation algorithm}\label{sec:algorithm}

The algorithm takes as input a complete Concept specification.
The output of the algorithm is a .sg file, a file representation of an FSM.
This can be imported into the \noun{Workcraft} tool-suite and viewed graphically as an FSM\@.

Psuedocode for the translation algorithm can be seen in Algorithm~\ref{alg:fsm}.
In brief, the algorithm begins by preparing the concept specification for the translation by combining the lists of arcs in a certain way (lines 1-5), 
and directly extracts pairs of causes and effects from the Concepts specification (lines 6--10).
It then infers all possible source and destination states with associated arcs (lines 11--22).

For our discussion of the algorithm, we will use the C-Element with environment 
example, the translated FSM and STG for which can be seen in 
Figure~\ref{fig:c-element-env}. Figure~\ref{fig:concepts_file} contains the 
concepts file for this example, which can be translated to either an FSM or STG. 

\begin{figure}[H]
	\centering
		\begin{minted}[fontsize=\normalsize]{haskell}
        	  module CElement where
        	  import Tuura.Concept.FSM

        	  circuit a b c = protocol <> ports <> init 
            	  where
                	  protocol = cElement a b c <> env
                	  env      = inverter c a <> inverter c b
                	  ports    = inputs[a, b] <> outputs[c] 
                	  init     = initialise0 [a, b, c]
		\end{minted}
	\vspace{-5mm}
	\begin{centering}
		\protect\caption{\label{fig:concepts_file}A concepts file for C-Element with an environment}
		\par\end{centering}
	
\end{figure}

There can be multiple ways of representing a specification using concepts, 
however, all levels of abstraction available to the designer are built out of signal-level
concepts. Given a specification, such as that shown in Figure~\ref{fig:concepts_file}, we can 
therefore break down 
all gate- and protocol-level constructs into `atoms', which significantly 
simplifies the translation task. For now, we will ignore initial states and interface
concepts, and focus on the arcs only.

\begin{table}[h]
\vspace{-2mm}
\caption{Lists of cause transitions by effect transitions
		\label{tab:list-of-concepts}}
  \centering
\begin{tabular}{@{}ll@{}}
	\toprule
	Causes & Effect\\ \midrule
	$a^{+}$, $b^{+}$	& $c^{+}$ \\
	$a^{-}$, $b^{-}$ 	& $c^{-}$ \\
	$c^{-}$			& $a^{+}$ \\
	$c^{+}$			& $a^{-}$ \\
	$c^{-}$			& $b^{+}$ \\
	$c^{+}$			& $b^{-}$ \\
	\bottomrule
\end{tabular}
\vspace{-5mm}
\end{table}

To start with we list all causes for each effect. 
These lists can be found in Table~\ref{tab:list-of-concepts}.
All of the included causalities are AND-causalities, and hence, there
is at least one single cause transitions for each effect transition. 

If OR-causality was included, the list of possible causes would be included with these
AND-causalities. For example:  $a^{+}$, $b^{+}$, $[x^{+}, y^{+}]$ causing
$c^{+}$. $x$ and $y$ are in a list of their own, and this shows OR-causality;
either $x$ or $y$ must transition high for $c^{+}$ to occur. However, this does
not account for the other AND-causalities. 

In this case, we perform the \emph{Cartesian product}, which combines
all AND- and OR-causalities. For the example above, this would produce:

\vspace{-2mm}

\begin{center}

$a^{+}$, $b^{+}$, $x^{+}$ causes $c/1^{+}$

$a^{+}$, $b^{+}$, $y^{+}$ causes $c/2^{+}$

\end{center}

\vspace{-2mm}

\noindent This includes one of the OR-causality transitions with all of the 
AND-causality transitions. It also leaves us with two separate $c^{+}$ 
transitions. These are both needed in order to show the two possible 
combinations of transitions necessary for $c^{+}$, and would be included as two
separate transition objects in an STG. 

However, for the C-element with environment example, there is no OR-causality. 
We still perform the step of applying the Cartesian product to each of these 
lists of transitions for each effect transition, but the result will be the 
same as the original lists, seen in Table~\ref{tab:list-of-concepts}.

In order to create a complete encoding of a single state, we must first find all of the signals that exist in the system.
We do this by first removing any \emph{effect} transition which may occur in the \emph{cause} list to avoid duplicate signals.
We then construct a signal list from each causality.
Taking the union of the signal lists from all causalities then gives us the complete set of signals in the system.
In our example, we obtain the list of signals $a$, $b$, $c$.

Since, for the specific type of FSM that the algorithm translates the specification to, FSTs, each state must contain information for every signal, 
and we must add to each causality any signals that are missing.
\Cref{tab:enc} shows the new signals in bold, which are added with an `x', or `don't care' polarity since both rising or falling transitions give valid states for these signals.

Now that all signals in the design are accounted for in each causality (that is, the concatenation of \emph{Causes} and \emph{Effect} in \cref{tab:enc} contains all the signals \texttt{a}, \texttt{b}, and \texttt{c}), we can begin constructing state encodings.
We sort the signals such that the binary encoding will represent ascending signal names from left to right -- in this example, \texttt{abc}.
Since the \emph{effect} transition tells us what the effect signal \emph{will become}, we have already obtained the destination state encodings for each arc in the FSM\@.
To obtain the source state encodings, we must flip the polarity of the \emph{effect} transition.
The result is shown in \cref{tab:enc}.

Note that we still have `don't care' polarity in some of our state encodings.
Since this means the corresponding signal can be either high or low, we must translate such states into all possible states to obtain a complete model\@.
We expand these don't care states into all of the possible congruent transition permutations.
For example, from the final row in \cref{tab:enc} we have the states \texttt{x01}, \texttt{x11}.
These will be expanded to the state pairs \texttt{001}, \texttt{011} and \texttt{101}, \texttt{111}.
\Cref{tab:enc_complete} gives the complete source and destination state encodings.
The \emph{effect} transition becomes the arc which links the source state to the destination state in the FSM\@.

\begin{table}[h]
\centering
%
\hfill
%
\parbox{0.45\linewidth}{%
\caption{Missing signals added to cause list (bold) with incomplete state encoding.}\label{tab:enc}
\begin{tabular}{@{}llll@{}}
	\toprule
	& & \multicolumn{2}{c}{Encoding (\texttt{abc})}\\ \cmidrule(l){3-4}
	Causes & Effect & Destination & Source\\ \midrule
	$a^{+}$, $b^{+}$		& $c^{+}$ & \texttt{111} & \texttt{110}\\
	$a^{-}$, $b^{-}$ 		& $c^{-}$ & \texttt{000} & \texttt{001}\\
	$\mathbf{b^{x}}$, $c^{-}$	& $a^{+}$ & \texttt{1x0} & \texttt{0x0}\\
	$\mathbf{b^{x}}$, $c^{+}$	& $a^{-}$ & \texttt{0x1} & \texttt{1x1}\\
	$\mathbf{a^{x}}$, $c^{-}$	& $b^{+}$ & \texttt{x10} & \texttt{x00}\\
	$\mathbf{a^{x}}$, $c^{+}$	& $b^{-}$ & \texttt{x01} & \texttt{x11}\\
	\bottomrule
\end{tabular}
}
%
\hfill
%
\parbox{0.45\linewidth}{%
\caption{Complete state encodings after expansion of `don't care' states.}\label{tab:enc_complete}
\begin{tabular}{@{}llll@{}}
	\toprule
	& \multicolumn{2}{c}{Encoding (\texttt{abc})}\\ \cmidrule(l){2-3}
	Effect & Destination & Source\\ \midrule
	$c^{+}$ & \texttt{111} & \texttt{110}\\[0.25em]
	$c^{-}$ & \texttt{000} & \texttt{001}\\[0.25em]
	$a^{+}$ & \texttt{100} & \texttt{000}\\
	$a^{+}$ & \texttt{110} & \texttt{010}\\[0.25em]
	$a^{-}$ & \texttt{001} & \texttt{101}\\
	$a^{-}$ & \texttt{011} & \texttt{111}\\[0.25em]
	$b^{+}$ & \texttt{010} & \texttt{000}\\
	$b^{+}$ & \texttt{110} & \texttt{100}\\[0.25em]
	$b^{-}$ & \texttt{001} & \texttt{011}\\
	$b^{-}$ & \texttt{101} & \texttt{111}\\
	\bottomrule
\end{tabular}
}
\end{table}

\begin{algorithm}[H]
\begin{algorithmic}[1]	% Optional param is frequency of line numbering
	\caption{Concepts to FSM Translation Algorithm\label{alg:fsm}}
	\For {\textbf{each} effect $e$}
                    \State allCauses $\leftarrow$ \textbf{Concatenate} lists of causes for $e$ \label{alg:list}
                    \State [causes] $\leftarrow$ \textbf{CartesianProduct} allCauses
                    \State \textbf{addAll} $causalitity$ ([causes], effect) \textbf{to} causalities
          \EndFor
	\For {\textbf{each} \emph{causality} ([cause], effect) \textbf{in} causalities}
		\State \textbf{removeAll} effect \textbf{from} [cause]
		\State [signalList] $\leftarrow$ \textbf{concatenate} ([cause], effect)
	\EndFor
	\State allSignals $\leftarrow$ \textbf{setUnion} ([[signalList]])
	\For {\textbf{each} \emph{transitionList} tl \textbf{in} [[signalList]]}
		\State missingTransitions $\leftarrow$ \textbf{setDifference} (tl, allSignals)
		\For {\textbf{each} \emph{transition} t \textbf{in} missingTransitions}
			\State \textbf{setPolarity} (t, `x')
		\EndFor
		\State allTransitions $\leftarrow$ \textbf{concatenate} (tl, missingTransitions)
		\State sortedTransitions $\leftarrow$ \textbf{sortBySignalName} (allTransitions)
		\State destEncodings $\leftarrow$ \textbf{getPolarity} (sortedTransitions)
		\State srcEncodings $\leftarrow$ \textbf{flipEffectPolarity} (destEncodings)
		\State incompleteArcs $\leftarrow$ \textbf{makeArcs} (srcEncodings, destEncodings)
		\State \textbf{expandX} (incompleteArcs)
	\EndFor
\end{algorithmic}
\end{algorithm}
